---
title: "Data Cleaning and Preliminary Analyses"
output: html_notebook
---

This document is where I will perform preliminary analyses to investigate my data. I will also conduct data cleaning necessary to conduct these analyses. 

Outline:

Data Cleaning:
0. Load in the data and necessary packages
1. Clean the data 
2. Calculate necessary parameters (development time in days not dates, ln(mass))
3. Get data into long format for making cumulative growth curve graphs

Survival Analyses: 
How many caterpillars survived the experiment? Just parasitized individuals. 

Growth Analyses:
4. Run models for mass gain and development time with diet and parasitization status as fixed effects, and bin as a random effect
5. Graph growth curves (ln-transformed) for total development
6. Plot mass and development time at wandering raw data, to map onto Geoff and I's study predictions

Wasp Analyses:
7. Run models for wasp survival to eclosion with diet, number of times parasitized, and wetness as fixed effects, and bin as a random effect
8. Graph boxplots for eclosion percentage across diet type 

# Data Cleaning

## 0. Load in the data and necessary packages

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lme4)
library(lubridate)
library(tidyr)
library(Rmisc)
library(lmerTest)
library(nlme)
```

```{r}
data <- read.csv("~/Desktop/MsCcSurvJun22.csv", header = TRUE)
```

## 1. Clean the data

### Eliminate those that died:

```{r}
data <- data[data$survived == "yes", ]
```

### Eliminate failed parasitizations:

```{r}

data %>%
  filter((data$para == "yes" & data$mass.w != "NA"))

data <- data[data$ID != 34, ]
data <- data[data$ID != 36, ]
data <- data[data$ID != 102, ]
data <- data[data$ID != 124, ]

```


### Checking sample sizes

```{r}
count(data[data$food == "DC" & data$para == "yes", ])
count(data[data$food == "DC" & data$para == "no", ])
count(data[data$food == "TB" & data$para == "yes", ])
count(data[data$food == "TB" & data$para == "no", ])
count(data[data$food == "diet" & data$para == "yes", ])
count(data[data$food == "diet" & data$para == "no", ])
```

## 2. Calculate necessary parameters

### Convert dates to experiment days 

```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}

# Run the loop for the various dates listed

data$day.hatch <- loop(start.date = "04/21/2020", date.list = data$date.hatch)
data$day.3rd <- loop(start.date = "04/21/2020", date.list = data$date.3rd)
data$day.para <- loop(start.date = "04/21/2020", date.list = data$date.para)
data$day.4th <- loop(start.date = "04/21/2020", date.list = data$date.4th)
data$day.5th <- loop(start.date = "04/21/2020", date.list = data$date.5th)
data$day.w <- loop(start.date = "04/21/2020", date.list = data$date.w)
data$day.pupa <- loop(start.date = "04/21/2020", date.list = data$date.pupa)
data$day.mass.pupa <- loop(start.date = "04/21/2020", date.list = data$date.mass.pupa)
data$day.eclos <- loop(start.date = "04/21/2020", date.list = data$date.eclos)
data$day.coc <- loop(start.date = "04/21/2020", date.list = data$date.coc)
data$day.remove <- loop(start.date = "04/21/2020", date.list = data$date.remove)
data$day.w.eclos <- loop(start.date = "04/21/2020", date.list = data$date.w.eclos)

```

### Calculate development times

```{r}
# Time til each instar
data$tth <- 0
data$tt3 <- data$day.3rd - data$day.hatch
data$tt4 <- data$day.4th - data$day.hatch
data$tt5 <- data$day.5th - data$day.hatch
data$ttw <- data$day.w - data$day.hatch
data$ttp <- data$day.pupa - data$day.hatch
data$tte <- data$day.eclos - data$day.hatch
data$ttcoc <- data$day.coc - data$day.hatch
data$ttweclos <- data$day.w.eclos - data$day.hatch

# Time spent in each instar/stage
data$delta3 <- data$day.4th - data$day.3rd
data$delta4 <- data$day.5th - data$day.4th
data$delta5 <- data$day.w - data$day.5th
data$deltap <- data$day.eclos - data$day.pupa

```

### Convert masses to ln(mass)

```{r}
data$lnh <- log(data$mass.hatch)
data$ln3 <- log(data$mass.3rd)
data$ln4 <- log(data$mass.4th)
data$ln5 <- log(data$mass.5th)
data$lnw <- log(data$mass.w)
```

### Calculate total number of wasps emerged and percent wasps that spun cocoons

```{r}
data$num.w.tot <- data$num.w.coc + data$num.w.larvae
data$perc.w.coc <- data$num.w.coc / data$num.w.tot
```


## 3. Convert to long format for graphing

### Converting to long format

```{r}
masslong <- data %>% gather(instar, mass, lnh, ln3, ln4, ln5, lnw)
masslong$instar <- gsub("ln", "", masslong$instar)

agelong <- data %>% gather(instar, age, tth, tt3, tt4, tt5, ttw)
agelong$instar <- gsub("tt", "", agelong$instar)

masslong <- masslong %>% select(ID, food, para, instar, mass)
agelong <- agelong %>% select(ID, food, para, instar, age)

datalong<-merge(masslong, agelong, by=c("ID", "food", "para", "instar"))

```

### Finding means (not separated by para status) 

```{r}
#calculate means and variation

#constructing summarySE for age and mass to make mean plot

masssum <- summarySE(datalong, measurevar = "mass",
                  groupvars = c("food", "instar"),
                  na.rm=TRUE)

agesum <- summarySE(datalong, measurevar = "age",
                   groupvars = c("food", "instar"),
                   na.rm = TRUE)

#combine into one data frame

means <- masssum

means$age <- agesum[, 4]
means$age.se <- agesum[, 6]
means$age.ci <- agesum[, 7]

```

### Finding means (separated by para status) 

```{r}
#calculate means and variation

#constructing summarySE for age and mass to make mean plot

pmasssum <- summarySE(datalong, measurevar = "mass",
                  groupvars = c("food", "para", "instar"),
                  na.rm=TRUE)

pagesum <- summarySE(datalong, measurevar = "age",
                   groupvars = c("food", "para", "instar"),
                   na.rm = TRUE)

#combine into one data frame

pmeans <- pmasssum

pmeans$age <- pagesum[, 5]
pmeans$age.se <- pagesum[, 7]
pmeans$age.ci <- pagesum[, 8]

```

# Survival Analyses:

## How many caterpillars survived parasitization?

```{r}
## Reload for those that didn't survive also 

survdata <- read.csv("~/Desktop/MsCcSurvJun22.csv", header = TRUE)
survdata$bin <- gsub("0", "", survdata$bin)
survdata$survived <- gsub("yes", "1", survdata$survived)
survdata$survived <- gsub("no", "0", survdata$survived)
survdata$survived <- as.factor(survdata$survived)

surv1 <- glmer(survived ~ food + (1|bin), 
            data = survdata[survdata$para == "yes", ], family = "binomial")

summary(surv1)
```

```{r}
surv2 <- glmer(survived ~ food + para + (1|bin), 
            data = survdata, family = "binomial")

summary(surv2)

#anova(surv2)
```

```{r}
surv3 <- glmer(survived ~ food + para + food:para + (1|bin), 
            data = survdata, family = "binomial")

summary(surv3)
```

```{r}
ggplot(aes(x = food, y = survived, color = para), data = data) +
  geom_point(alpha = 0.8, position = position_jitter(width = 0.2, height = .2)) + 
  scale_color_manual(values=c("#95D840", "#440D54"),
                     name="Parasitization status",
                     label=c("Unparasitized", "Parasitized")) + 
  labs(x = "Diet Type", y = "Survival")
```


# Growth Analyses:

## 4. Run models for mass gain and development time with diet as a fixed effect, and bin as a random effect

### Change 0s to NAs for bin on diet individuals

```{r}
data$bin <- gsub("0", "", data$bin)
```

### Checking relationship between day massed and weight

```{r}
ggplot(aes(x = (day.mass.pupa - day.pupa), y = mass.pupa, color = food), data = data[data$para == "no", ]) + 
  geom_jitter(alpha = 0.8, width = 0.2) +  
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco"))
  
```

No obvious relationship - I don't think we need to account for it in our model. Let's run a model to make sure though:

```{r}
data$for.check <- data$day.mass.pupa - data$day.pupa

check1 <- lmer(mass.pupa ~ for.check + (1|bin), data = data[data$para == "no", ])

anova(check1)
```

Great! No relationship. 

### Mass at eclosion

```{r}
me1 <- lmer(mass.eclos ~ food + sex + food:sex + (1|bin), data = data[data$para == "no",])

anova(me1)
```

```{r}
ggplot(aes(x = food, y = mass.eclos), data = data) + 
  geom_boxplot()
```

### Time as a pupa

```{r}
tp1 <- lmer(deltap ~ food + sex + food:sex + (1|bin), data = data[data$para == "no", ])

anova(tp1)
```

```{r}
data.graph <- data[complete.cases(data$sex), ]

ggplot(aes(x = food, y = deltap, color = sex), data = data[data$para == "no", ]) + 
  geom_point(alpha = 0.8, position = position_jitter(width = 0.1)) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Sex",
                     label=c("NA", "Female", "Male")) + 
  labs(x = "Diet Type", y = "Time spent as a pupa (days)")
```


### Mass at pupation 

```{r}
mp1 <- lmer(mass.pupa ~ food + sex + food:sex + (1|bin), data = data[data$para == "no", ])

anova(mp1)

```

Food is highly significant - sex is not. Let's plot this:

```{r}
ggplot(aes(x = food, y = mass.pupa), data = data[data$para == "no", ]) + 
  geom_boxplot() + 
  labs(x = "Diet Type", y = "Mass of pupa (mg)")
  
```


### Mass at wandering 

```{r}
mw1 <- lmer(mass.w ~ food + (1|bin), data = data[data$para == "no", ])

anova(mw1)

```

```{r}
ggplot(aes(x = food, y = mass.w), data = data[data$para == "no", ]) + 
  geom_boxplot() + 
  labs(x = "Diet Type", y = "Mass at wandering (mg)")
  
```

Compared to the pupation data, this means that DC bugs lost more mass during pupation than the other two groups, and TB lost more than those on diet. Very interesting! 

```{r}
mw2 <- lmer(mass.w ~ food + (1|bin), data = data[data$para == "no" & data$food != "diet", ])

summary(mw2)
anova(mw2)

```

No interaction of diet overall - is an effect when you remove the diet category

### Calculating mass lost

```{r}
data$mass.lost.pupa <- data$mass.w - data$mass.pupa
```

Look at average values:

```{r}
tbml <- data[data$food == "TB" & data$para == "no", ]
dcml <- data[data$food == "DC" & data$para == "no" & data$ID != 29, ]
dietml <- data[data$food == "diet" & data$para == "no", ]

mean(tbml$mass.lost.pupa)
mean(dcml$mass.lost.pupa)
mean(dietml$mass.lost.pupa)
```


Model real quick:

```{r}
ml1 <- lmer(mass.lost.pupa ~ food + sex + food:sex + (1|bin), data = data[data$para == "no", ])

#summary(ml1)

anova(ml1)

```

```{r}
ggplot(aes(x = food, y = mass.lost.pupa), data = data[data$para == "no", ]) + 
  geom_boxplot() + 
  labs(x = "Diet Type", y = "Mass lost during transition to pupa (mg)")
```

Diet is significant, not sex. 

### Time til wandering

```{r}
tw1 <- lmer(day.w ~ food + (1|bin), data = data[data$para == "no", ])

anova(tw1)

```

No effect of diet overall

### Mass at 5th instar

```{r}
m51 <- lmer(mass.5th ~ food + para +food:para + (1|bin), data = data)

anova(m51)
```

No interactions between para + food type (food*para)

### Time til 5th instar

```{r}
t51 <- lmer(day.5th ~ food + para + food:para + (1|bin), data = data)

anova(t51)
```

```{r}
t52 <- lmer(day.5th ~ para + (1|bin), data = data)

anova(t51, t52)
```

Also no food*para here 

Can't partition out effects of food vs food:para 
Marginal whether full model is significantly better

## 5. Graph growth curves (ln-transformed) for total development

### Combined with CIs

```{r}

meanplot <- ggplot(means, aes(x = age, y = mass, group = food, color = food)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.2) + 
  geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
                width = .3, size = 1) + 
  geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
                 height = .3, size = 1) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained ln(mg)")


ggsave("MsCcSurvJun16.png", width = 8, height = 5)

meanplot


```



### Separated para and non-para

```{r}

pmeanplot <- ggplot(pmeans, aes(x = age, y = mass, color = food, shape = para)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.2, aes(linetype = para)) + 
  geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
                width = .3, size = 1) + 
  geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
                 height = .3, size = 1) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  scale_shape_manual(values = c(19, 2)) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained ln(mg)") + 
  xlim(5, 17) + 
  ylim(3, 8.5)


ggsave("MsCcSurvParaJun16.png", width = 8, height = 5)

pmeanplot


```

Bigger effect on diet than on either host plant


## 6. Plot mass and development time at wandering raw data, to map onto Geoff and I's study predictions

```{r}
raw <- ggplot(data, aes(x = ttw, y = mass.w, color = food)) + 
  geom_point(alpha = 0.8, position = "jitter") + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Age at wandering (days)", y = "Mass at wandering (mg)")

ggsave("MsCcSurvRawJun16.png", width = 8, height = 5)

raw
```

## 7. Parasitoid Models

### Time til parasitoid emergence

```{r}
te1 <- lmer(day.coc ~ food + (1|bin), data = data[data$para == "yes", ])

anova(te1)
```

No effect on day of emergence (but it's singular?)

### Number of wasps emerged

```{r}
we1 <- lmer(num.w.tot ~ food + (1|bin), data = data[data$para == "yes", ])

anova(we1)

```

```{r}
we2 <- lmer(num.w.tot ~ food + num.para +food:num.para + (1|bin), data = data[data$para == "yes" & data$num.para != 3, ])

anova(we2)
```

```{r}
we3 <- lmer(num.w.tot ~ food + num.para + food:num.para + (1|bin), data = data[data$para == "yes" & data$num.para != 2, ])

anova(we3)
```



```{r}
data$num.para <- as.factor(data$num.para)

ggplot(aes(x = num.para, y = num.w.tot), data = data[data$para == "yes", ]) + 
  geom_boxplot()
```


No effect on number of wasps emerged

### Percent wasps that were cocoons - do this as a binomial! 

1. Get the data in binomial form 
2. Run the models 

```{r}
binom <- data[data$para == "yes", ] %>%
  select(ID, food, bin, num.para, num.w.coc, num.w.larvae, perc.w.coc)

```

```{r}
bin1 <- glmer(cbind(num.w.coc, num.w.larvae) ~ food + num.para + food:num.para + (1|bin), 
            data = binom, family = "binomial")

summary(bin1)

#anova(bin1)
```

```{r}
bin2 <- glmer(cbind(num.w.coc, num.w.larvae) ~ food + (1|bin), 
            data = binom, family = "binomial")

summary(bin2)
```

```{r}
anova(bin1, bin2)
```



```{r}
binom$num.para <- as.factor(binom$num.para)

ggplot(aes(x = food, y = perc.w.coc, color = num.para), data = binom) + 
  geom_point(alpha = 0.8, position = position_jitter(width = 0.1)) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Number of Parasitizations",
                     label=c("1", "2", "3")) + 
  labs(x = "Diet Type", y = "Percent of emerged wasps that spun cocoons")
```


Graph of # parasitizations below:

```{r}
ggplot(aes(x = num.para, y = perc.w.coc, color = food), data = data[data$para == "yes", ]) + 
  geom_jitter(width = 0.1, alpha = 0.8) + 
    scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Number of parasitizations", y = "Percentage of wasp larvae emerged that spun cocoons")

```

Without that one outlier in the 2 parasitizations group, I think this relationship would look a lot different.

### Percent wasps eclosed

### Sex ratio of wasps eclosed
