---
title: "Wasp Analyses"
output: html_notebook
---

This notebook is for cleaned analyses of wasp survival. 

Roadmap: 

0. Load in the data and packages
1. Clean dataset to desired columns and individuals (parasitized)
2. Investigate number of larvae emerged, percent of larvae eclosed, and proportion successful eclosions. 
3. Graph!

## 0. Load in the data and packages

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lme4)
library(lubridate)
library(tidyr)
library(Rmisc)
library(lmerTest)
library(nlme)
```

```{r}
data <- read.csv("~/Desktop/MsCcSurvNov5.csv", header = TRUE)
```


## 1. Clean data

Keep only surviving caterpillars:

```{r}
data <- data[data$survived == "yes", ]
```

Keep only parasitized caterpillars: 

```{r}
data <- data[data$para == "yes", ]
```

Remove failed parasitizations:

```{r}
data %>%
  filter((data$para == "yes" & data$mass.w != "NA"))

data <- data[data$ID != 34, ]
data <- data[data$ID != 36, ]
data <- data[data$ID != 102, ]
data <- data[data$ID != 124, ]

```

Keep ony columns of interest for parasitization investigation: 

```{r}
data <- data %>%
  select(ID, food, bin, num.para, mass.5th, num.w.coc, num.w.larvae, num.w.eclos.tot)
```

Make new columns for analyses:

```{r}
data$num.w.emerg.tot <- data$num.w.coc + data$num.w.larvae

data$perc.eclos <- data$num.w.eclos.tot / data$num.w.coc

data$surv.to.eclos <- data$num.w.eclos.tot / data$num.w.emerg.tot
```

## 2. Model data

```{r}
modadd <- lm(surv.to.eclos ~ food + mass.5th, data = data)

summary(modadd)
```
Nothing pops out as significant. 

```{r}
modint <- lm(surv.to.eclos ~ food*mass.5th, data = data)

summary(modint)
```
When you add in the interaction terms, now everything is significant... 

```{r}
modbin <- lmer(surv.to.eclos ~ food + mass.5th + (1|bin), data = data)

summary(modbin)
```

Singular when I add in bin - too many degrees of freedom I guess. 

## 3. Graph! 

```{r}
ggplot(aes(x = num.w.emerg.tot, y = surv.to.eclos, color = food, size = num.para), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"))
```

