---
title: "Wasp Analyses"
output: html_notebook
---

This notebook is for cleaned analyses of wasp survival. 

Roadmap: 

0. Load in the data and packages
1. Clean dataset to desired columns and individuals (parasitized)
2. Investigate number of larvae emerged, percent of larvae eclosed, and proportion successful eclosions. 
3. Graph!

## 0. Load in the data and packages

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lme4)
library(lubridate)
library(tidyr)
library(Rmisc)
library(lmerTest)
library(nlme)
```

```{r}
data <- read.csv("~/Desktop/GitHub/MsCcDietSurvival/Data/MsCcSurvDec31FINAL.csv", header = TRUE)
```


## 1. Clean data

Keep only surviving caterpillars:

```{r}
data <- data[data$survived == "yes", ]
```

Keep only parasitized caterpillars: 

```{r}
data <- data[data$para == "yes", ]
```

Remove failed parasitizations:

```{r}
data %>%
  filter((data$para == "yes" & data$mass.w != "NA"))

data <- data[data$ID != 34, ]
data <- data[data$ID != 36, ]
data <- data[data$ID != 102, ]
data <- data[data$ID != 124, ]

```

Keep ony columns of interest for parasitization investigation: 

```{r}
data <- data %>%
  select(ID, food, bin, num.para, mass.5th, num.w.coc, num.w.larvae, num.w.eclos.tot)
```

Make new columns for analyses:

```{r}
data$num.w.emerg.tot <- data$num.w.coc + data$num.w.larvae

data$perc.eclos <- data$num.w.eclos.tot / data$num.w.coc

data$surv.to.eclos <- data$num.w.eclos.tot / data$num.w.emerg.tot
```

## 2. Model data

```{r}
modadd <- lm(surv.to.eclos ~ food + mass.5th, data = data)

summary(modadd)
```
Nothing pops out as significant. 

```{r}
modint <- lm(surv.to.eclos ~ food*mass.5th, data = data)

summary(modint)
```
When you add in the interaction terms, now TB bugs are differentt than diet and DC bugs... 

```{r}
modbin <- lmer(surv.to.eclos ~ food + mass.5th + (1|bin), data = data)

summary(modbin)
```

Singular when I add in bin - too many degrees of freedom I guess.

### Analyzing within a binomial framework

```{r}
binom1 <- glm(surv.to.eclos ~ food, data = data, family = "binomial", weights = num.w.emerg.tot)

summary(binom1)

#binom1nw <- glm(surv.to.eclos ~ food, data = data, family = "binomial")

#summary(binom1nw)
```

Here, diet is popping out as being significantly different than DC and TB

```{r}
binom2 <- glm(surv.to.eclos ~ food*mass.5th, data = data, family = "binomial", weights = num.w.emerg.tot)

summary(binom2)
```

Now everything's significant - eek. 

```{r}
food <- lm(mass.5th ~ food, data = data)

anova(food)
```



## 3. Graph! 

```{r}
ggplot(aes(x = num.w.emerg.tot, y = surv.to.eclos, color = food), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Total number of wasp larvae emerged", y = "Percent survival to adulthood")
```


```{r}
ggplot(aes(x = num.w.emerg.tot, y = num.w.eclos.tot, color = food), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Total number of wasp larvae emerged", y = "Total number of wasp larvae eclosed")
```

```{r}
ggplot(aes(x = num.w.emerg.tot, y = num.w.coc, color = food), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Total number of wasp larvae emerged", y = "Total number of wasp cocoons")
```


```{r}
ggplot(aes(x = num.w.emerg.tot, y = surv.to.eclos, color = food, size = num.para), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Total number of wasp larvae emerged", y = "Percent survival to adulthood")
```

```{r}
ggplot(aes(x = food, y = perc.eclos), data = data) + 
  #geom_violin() + 
  geom_jitter(width = 0.05, alpha = 0.5) + 
  labs(x = "Diet Type", y = "Percent cocoons that successfully eclosed")
```

```{r}
ggplot(aes(x = num.w.emerg.tot, y = mass.5th, color = food), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Total number of wasp larvae emerged", y = "Mass at 5th instar")
```

I had a lot (17) with emergence during 4th instar - not sure how best to account for those, since I didn't measure final mass... Can I measure final mass while frozen?

```{r}
ggplot(aes(x = surv.to.eclos, y = mass.5th, color = food), data = data) + 
  geom_point() + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54")) + 
  labs(x = "Percent wasps survived to eclosion", y = "Mass at 5th instar")
```

New graph that Joel wanted: 

Make means for each stage

```{r}
means <- data %>%
  group_by(food) %>%
    summarise(emerged = mean(num.w.emerg.tot), cocoons = mean(num.w.coc), 
              eclosed = mean(num.w.eclos.tot))

meanslong <- gather(means, stage, number, emerged:eclosed)

meanslong$stage <- ordered(meanslong$stage, levels = c("emerged", "cocoons", "eclosed"))
```

```{r}
ggplot(aes(x = stage, y = number, color = food), data = meanslong) + 
  geom_point() + 
  geom_line(aes(group = food)) + 
  ylim(0, 110) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Life stage", y = "Mean number of individuals")
```

Need to get SEs in there somehow... 

```{r}
graphdata <- data %>%
  rename(emerged = num.w.emerg.tot) %>%
  rename(cocoons = num.w.coc) %>%
  rename(eclosed = num.w.eclos.tot)

graphdata <- graphdata %>%
  select(ID, food, emerged, cocoons, eclosed)

graphdatalong <- gather(graphdata, stage, number, emerged:eclosed)
graphdatalong$stage <- ordered(graphdatalong$stage, levels = c("emerged", "cocoons", "eclosed"))
```



ggplot(aes(x = stage, y = number, color = food), data = graphdatalong) + 
  geom_point(aes(alpha = 0.5)) + 
  geom_point(data = meanslong, aes(x = stage, y = number, color = food, size = 3)) + 
  geom_line(data = meanslong, aes(group = food)) + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Life stage", y = "Number of individuals") 


Get SE values: 

```{r}
ses <- summarySE(graphdatalong, measurevar = "number",
                  groupvars = c("food", "stage"),
                  na.rm=TRUE)
```

```{r}
ggplot(aes(x = stage, y = number, color = food), data = ses) + 
  geom_point(aes(size = 3)) + 
  geom_line(aes(group = food)) + 
  ylim(0, 120) + 
  geom_errorbar(aes(ymin = number - se, ymax = number + se),
                width = .1, size = .5) +
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  labs(x = "Life stage", y = "Mean number of individuals")
```

So this just doesn't match with my previous results... I think what's happening is that the binomial framework is analyzing from the difference between emerged and eclosed, not the raw eclosed values. Diet has the lowest dip between emerged and eclosed, meaning not that more diet survived in raw numbers, but that a higher proportion of diet survived. 
